---
title: "Problem Set 1"
author: "Luis Rivas"
subtitle: MGSC 310 Problem Set 1
output:
  html_document:
    df_print: paged
  html_notebook: default
---

```{r setup, include=FALSE}


library(knitr)

# As long as you are working in a Rstudio Project file, you shouldn't need to 'hard code' directories like this 
# change to your own working directory
#knitr::opts_knit$set(root.dir = 'C:/Users/doosti/Desktop/MGSC_310')
knitr::opts_knit$set(root.dir = '/Users/luisrivas/Documents/CPSC_Courses/MGSC_310/')

# set seed to your own favorite number
set.seed(310)
options(width=70)
# if you want to prevent scientific format for numbers use this:
options(scipen=99)

# general rchunk code options
opts_chunk$set(tidy.opts=list(width.wrap=50),tidy=FALSE, size = "vsmall")
opts_chunk$set(message = FALSE,
               warning = FALSE,
               cache = TRUE,
               autodep = TRUE,
               cache.comments = FALSE,
               collapse = TRUE,
               fig.width = 5,  
               fig.height = 4,
               fig.align='center')

```

```{r setup_2}

# load all your libraries here
library('tidyverse')
# note, do not run install.packages() inside a code chunk. install them in the console outside of a code chunk. 

```


## Question 1

1a) Checking that the working directory is set correctly. 

```{r}

# code for part a
getwd() 

```


1b) Reading IMDB_movies.csv dataset into movies. 

```{r}

# code for part b
movies <- read.csv("datasets/IMDB_movies.csv")

```


1c) This dataset has 3889 rows and 25 columns. 

```{r}

# code for 1c
dim(movies)

```


1d) These are the names of the variables in the movies dataset. 

```{r}

# code for 1d
names(movies)

```


1e) Here we can see the data type of each variable. The variables that are characters are: movie_title, director_name, country, language, genres, content_rating, color, plot_keywords, actor_1_name, actor_2_name. You can tell because they have "chr" after the variable name. 

```{r}

# code for 1e
movies %>% glimpse()

```


1f) Here are the first 5 rows of the dataset. 

```{r}

# code for 1f
movies %>% slice(1:5)

```


1g) Here are the top 10 grossing movies in the dataset in descending order. Director James Cameron is king of the box office. 

```{r}

# code for 1g
movies %>% arrange(-gross) %>% slice(1:10)

```


1h) Creating a smaller dataset called movies_sub which is simply the movies dataset but without the variable movies_sub. This new dataset still has the original 3889 rows but only has 24 columns, as opposed to movies which has 25 columns. 

```{r}

# code for 1h
movies_sub <- movies %>% select(-facenumber_in_poster)
dim(movies_sub)

```


1i) There are 1329 movies rated PG-13 in this dataset. The average gross for PG-13 movies in this dataset is $68,270,240. 

```{r}

# code for 1i
pg13 <- movies %>% filter(content_rating == "PG-13") 
nrow(pg13) 
mean(pg13$gross)

```



## Question 2

2a) Scatter plot of IMDB score on the x-axis and gross on the y-axis. 


```{r}

# code for 2a
library(ggplot2)
ggplot(movies, aes(x = imdb_score, y = gross)) + geom_point()

```


2b) Same scatter plot but with reduced transparency of each point. 


```{r}

# code for 2b
ggplot(movies, aes(x = imdb_score, y = gross)) + geom_point(alpha = 1/10)

```


2c) Same scatter plot but with y-axis in millions of dollars. 


```{r}

# code for 2c
movies <- movies %>% mutate(grossM = gross/1000000)
ggplot(movies, aes(x = imdb_score, y = grossM)) + geom_point(alpha = 1/10)

```


2d) Same scatter plot but with a smoothing line. The relationship between movie gross and IMDB score is positive. In general, when one goes up, the other also goes up. 


```{r}

# code for 2d
ggplot(movies, aes(x = imdb_score, y = grossM)) + geom_point(alpha = 1/10) + geom_smooth()

```


2e) Scatter plot showing movies gross against IMDB score for James Cameron films, where the points are the movie titles. 


```{r}

# code for 2e
ggplot(movies %>% filter(director_name == "James Cameron"), aes(x = imdb_score, y = grossM)) + geom_text(aes(label = movie_title), color = "blue") 

```


2f) Same scatter plot but with more reasonable ranges for the axes. 


```{r}

# code for 2f
ggplot(movies %>% filter(director_name == "James Cameron"), aes(x = imdb_score, y = grossM)) + geom_text(aes(label = movie_title), color = "blue") + xlim(7,9) + ylim(0,800)

```


2g) Same scatter plot but with better axes titles. 


```{r}

# code for 2g
ggplot(movies %>% filter(director_name == "James Cameron"), aes(x = imdb_score, y = grossM)) + 
  geom_text(aes(label = movie_title), color = "blue") + xlim(7,9) + ylim(0,800) + 
  xlab("Movie IMDB Score") + ylab("Movie Gross in Millions")

```


